\Large \bf One-Step, Three-Factor Authentication With Custom-Fit, In-Ear EEG

### 2017-02-14 paper.tex
Missed a table for bolding
Adding section in discussion about sample size

\section{Related work}

@@

\subsection{Usable authentication}

@@

\subsection{Passthought authentication}
The use of EEG as a biometric signal for user authentication has a short history.
In 2005, Thorpe et al. motivate and outline the design of a passthoughts system,
where, rather than typing a password, users authenticate 
by thinking of a passthought \cite{Thorpe2005}.
Since 2002, a number of independent groups have achieved 99-
100\% authentication accuracy using multi-channel sensors placed on the scalp \cite{Poulos2002,Marcel2007a,Palaniappan2008,Ashby2011}.
In 2013, one group showed that 99\% authentication accuracy can also be
achieved using a consumer-grade single-channel sensor \cite{Chuang2013b}. In particular, the
lack of signal diversity from multiple EEG channels can be overcome by allowing
the users to choose their own personalized passthoughts (e.g., sing their favorite
song in their head). There are two significant consequences of this result. First,
the passthoughts approach is no longer constrained by the high cost (\textgreater \$10,000 USD)
and low usability (gel-based electrodes; aesthetic challenges of an EEG cap) of
medical-grade multi-channel devices. Second, because users can choose and
easily change their secret mental task, this approach can support one-step two-
factor authentication via the simultaneous presentation of the inherence factor
(brainwave signatures due to the unique folding structures of the cortex) and the
knowledge factor (the secret mental task) \cite{Chuang2014}.

@@

\subsection{Manufacturing and materials}

@@

The first quality check was examining the electrical impedances achieved for the ear pieces as a low impedance (<10 kOhm) 
is best for obtaining quality EEG signals. Table \ref{tab:impedances} below summarizes the impedances across the seven participants' six ear channels. With the exception of a few channels in select participants, impedances achieved were overall good.

@@

\subsection{Authentication results in the left-ear}

In the previous section, we trained and tested a passthought classifier with each task as the passthought, for all participants. Focusing on the left ear, we filter our results for the best-performing tasks. We rank tasks by lowest FAR and, given a tie, the lowest FRR (Table \ref{tab:best}).

@@

A subset of four participants completed our usability questionnaire. This questionnaire asked them to rank each of the mental tasks on 7-point likert-type scales on ease of use, level of engagement, perceived repeatability, and likeliness to use in a real-world authentication setting. The breathe and listen (\(\mu\)=6.75) tasks were ranked easiest to use, while the sequence (\(\mu\)=2.25) and face (\(\mu\)=2.75) tasks were ranked the most difficult. Sequence was rated highly in engagement however (\(\mu\)=5), as was the song task (\(\mu\)=5), while breathe-open and listen were ranked least engaging (\(\mu\)=2.25). Breathe (\(\mu\)=7) and listen (\(\mu\)=6.75) were ranked highest in repeatability, with sequence (\(\mu\)=2.5) along with face and sport (\(\mu\)=3) ranked least repeatable. Finally in terms of likeliness of use for authentication, song-open (\(\mu\)=5) and sequence (\(\mu\)=4.25) performed highest though modestly, while breathe (\(\mu\)=2.75) and listen (\(\mu\)=3) were rated least likely to use. In addition to these rating scales we asked participants to rank the tasks from most (1) to least (9) favorite overall. The song-open task ranked highest among favorites (\(\mu\)=4.25) followed by a tie between breathe-open, song, and speech (\(\mu\)=4.75). In line with some of the results above, sequence (\(\mu\)=7.75) and face (\(\mu\)=6.75) were rated as least favorites.

@@

Our findings demonstrate the apparent feasibility of single earpiece, achieving good results with only three electrodes and a reference, all on the left ear. FARs and FRRs are low across all participants and tasks, with FARs overall lower than FRRs. Participants' best-performing passthoughts typically seeing no errors in our training. Furthermore, no spoofed attacks were successful in our cursory analysis.

The powerful interactions between inherence and knowledge emerged in our spoofing attack. Although our target participant documented their chosen passthought, the spoofers found ambiguity in how these passthoughts could be expressed. For the face task, the spoofers did not know the friend the original participant had chosen. For the song tasks, though the song was known, the spoofers did not know what part of the song the original participant had imagined, or how it was imagined (humming, imagining a full performance, melody, vocals, etc). This experience sheds light on the highly individual nature of passthoughts, and provides a positive indication that there may be some intrinsic difficulty ofr spoofing passthoughts.

In our analysis, some notable patterns emerged. First, \textit{breathe} tended to be the best-performing task among participants. Classifiers overall distinguished the breath task even compared to breath tasks from other participants, implying that the task is expressed differently for each participant, i.e. that this task has an inherence factor sufficient for authentication, even though the task does not have a knowledge factor. Second, we were able to achieve good results by generating feature vectors based on only 500ms (300 voltage readings across the three electrodes). This short timespan is somewhat surprising, given that some tasks (like songs) presumably rely on changes or patterns over a longer period of time.

Furthermore, the use of conductive gel results in low impedances on other participants' custom-fit earpieces, despite the uniquness of ear canal shapes between individuals \cite{Akkermans2005}.
Nevertheless, classifiers appear to resists spoofing attacks, indicating that task-related signals are unique to individuals.
In the future, the custom-fit earbuds could include a hardware keypair to sign authentication attempts, providing a more secure method for establishing the posession factor in our three-factor, one-step scheme.

Finally, performance on Fp1 was not as high as performance in the ear, despite Fp1's popularity in past work on passthoughts \cite{Chuang2013b}. This could be explained by the greater number of electrodes in the ear (compared to just one on Fp1). Additionally, Fp1 is best poised to pick up on frontal lobe activity (e..g, concentration), but our tasks did not generally involve frontal lobe activity; in fact, a good number of them involved audio, which we would expect to be better observed near the auditory cortex at the ears. Future work should continue to investigate what sorts of mental tasks lend themselves to in-ear recording.

>>>>

\section{Related Work}

@@

\subsection{Usable Authentication}

@@

\subsection{Passthoughts Authentication}
The use of EEG as a biometric signal for user authentication has a short history.
In 2005, Thorpe et al. motivate and outline the design of a passthoughts system,
where, rather than typing a password, users authenticate 
by thinking of a passthought \cite{Thorpe2005}.
Since 2002, a number of independent groups have achieved 99-
100\% authentication accuracy using multi-channel sensors placed on the scalp \cite{Poulos2002,Marcel2007a,Palaniappan2008,Ashby2011}.
In 2013, one group showed that 99\% authentication accuracy can also be
achieved using a consumer-grade single-channel sensor \cite{Chuang2013b}. In particular, the
lack of signal diversity from multiple EEG channels can be overcome by allowing
the users to choose their own personalized passthoughts (e.g., sing their favorite
song in their head). There are two significant consequences of this result. First,
the passthoughts approach is no longer constrained by the high cost (\textgreater \$10,000 USD)
and low usability (gel-based electrodes; aesthetic challenges of an EEG cap) of
medical-grade multi-channel devices. Second, because users can choose and
easily change their secret mental task, this approach can support one-step two-
factor authentication via the simultaneous presentation of the inherence factor
(brainwave signatures due to the unique folding structures of the cortex) and the
knowledge factor (the secret mental task) \cite{Chuang2014}.

@@

\subsection{Earpiece Design and Manufacturing}

@@

The first quality check was examining the electrical impedances achieved for the ear pieces as a low impedance (\textless10 kOhm) 
is best for obtaining quality EEG signals. Table \ref{tab:impedances} below summarizes the impedances across the seven participants' six ear channels. With the exception of a few channels in select participants, impedances achieved were overall good.

@@

\subsection{Authentication Results}

In the previous section, we trained and tested a passthought classifier with each task as the passthought, for all participants. Our analyses revealed the choice between left and right ear to have little effect on the results, this enables us to postulate around a more usable device using channels only around the left ear.  Focusing on the left ear, we filter our results for the best-performing tasks. We rank tasks by lowest FAR and, given a tie, the lowest FRR (Table \ref{tab:best}).

@@

A subset of four participants completed our usability questionnaire. This questionnaire asked them to rank each of the mental tasks on 7-point likert-type scales on ease of use, level of engagement, perceived repeatability, and likeliness to use in a real-world authentication setting. The breathe and listen (\(\mu\)=6.75) tasks were ranked easiest to use, while the sequence (\(\mu\)=2.25) and face (\(\mu\)=2.75) tasks were ranked the most difficult. Sequence was rated highly in engagement however (\(\mu\)=5), as was the song task (\(\mu\)=5), while breathe-open and listen were ranked least engaging (\(\mu\)=2.25). Breathe (\(\mu\)=7) and listen (\(\mu\)=6.75) were ranked highest in repeatability, with sequence (\(\mu\)=2.5) along with face and sport (\(\mu\)=3) ranked least repeatable. Finally in terms of likeliness of use for authentication, song-open (\(\mu\)=5) and sequence (\(\mu\)=4.25) performed highest though modestly, while breathe (\(\mu\)=2.75) and listen (\(\mu\)=3) were rated least likely to use.

In addition to these rating scales we asked participants to rank the tasks from most (1) to least (9) favorite overall. The song-open task ranked highest among favorites (\(\mu\)=4.25) followed by a tie between breathe-open, song, and speech (\(\mu\)=4.75). In line with some of the results above, sequence (\(\mu\)=7.75) and face (\(\mu\)=6.75) were rated as least favorites.

@@

Our findings demonstrate the apparent feasibility of single earpiece, achieving good results with only three electrodes and a reference, all on or around the left ear. FARs and FRRs are low across all participants and tasks, with FARs overall lower than FRRs, a desirable pattern in terms of authenticating access to potentially sensitive information. Participants' best-performing passthoughts typically see no errors in our training. Furthermore, no spoofed attacks were successful in our cursory analysis.

This study was conducted with a small, relatively homogeneous sample of participants. We recognize this weakness and agree that more participants from diverse sample would greatly improve our understanding and interpretation of the capabilities of this technology. This does not lessen the importance of the work reported here however, as this sample size is comparable to that of previous scalp-based EEG passthoughts work\cite{Ashby2011, Marcel2007a, Palaniappan2008, palaniappan2006electroencephalogram, Poulos2002, Chuang2013b}, the only other in-ear EEG passthoughts study we are aware of \cite{curranpassthoughts} (note a generic fit system was used here), and other custom-fit in-ear EEG research \cite{kidmose2013study}. The fitting and manufacturing of custom-fit earpieces was the main limitation to increasing our sample size, and may very well pose a limitation to adoption of such a technology. Recently, however there have been developments in at-home kits for creating one's own custom-fitted earpieces \cite{voix2015settable} that could help overcome this barrier.

The powerful interactions between inherence and knowledge emerged in our spoofing attack. Although our target participant documented their chosen passthought, the spoofers found ambiguity in how these passthoughts could be expressed. For the face task, the spoofers did not know the friend the original participant had chosen. For the song tasks, though the song was known, the spoofers did not know what part of the song the original participant had imagined, or how it was imagined (humming, imagining a full performance, melody, vocals, etc). This experience sheds light on the highly individual nature of passthoughts, and provides a positive indication that there may be some intrinsic difficulty of spoofing passthoughts.

In our analysis, some notable patterns emerged. First, \textit{breathe} tended to be the best-performing task among participants. Classifiers overall distinguished the breathe task even compared to breathe tasks from other participants, implying that the task is expressed differently for each participant, i.e. that this task has an inherence factor sufficient for authentication, even though the task does not have a secondary knowledge factor beyond knowing which task to perform. Second, we were able to achieve good results by generating feature vectors based on only 500ms (300 voltage readings across the three electrodes). This short timespan is somewhat surprising, given that some tasks (like songs) presumably rely on changes or patterns over a longer period of time.

Furthermore, the use of conductive gel results in low impedances on other participants' custom-fit earpieces, despite the uniqueness of ear canal shapes between individuals \cite{Akkermans2005}.
Nevertheless, classifiers appear to resists spoofing attacks, indicating that task-related signals are unique to individuals.
In the future, the custom-fit earbuds could easily include a hardware keypair to sign authentication attempts, providing a secure method for establishing the possession factor in our three-factor, one-step scheme.

Finally, performance on Fp1 was not as high as performance in the ear, despite Fp1's popularity in past work on passthoughts \cite{Chuang2013b} and the general assumed benefits of measuring EEG from scalp-based locations. This could be explained by the greater number of electrodes in the ear (compared to just one on Fp1). Additionally, Fp1 is best poised to pick up on frontal lobe activity (e..g, concentration), but our tasks did not generally involve frontal lobe activity; in fact, a good number of them involved audio, which we would expect to be better observed near the auditory cortex at the ears. Future work should continue to investigate what sorts of mental tasks lend themselves to in-ear recording.
